Top Mortality causes in young adults and middle age adults in Switzerland

### 2024-04-21 paper/paper.qmd
update Data section

---
title: "Top Mortality causes in young adults and middle age adults in Switzerland"
subtitle: "Analyzing Switzerland Mortality causes data in 2010, 2015, adn 2019 from WHO database"
author: 
  - Thu Dong
thanks: "Code and data are available at: https://github.com/thu2912/Top-mortality-causes-of-age-groups-in-Switzerland.git"
date: today
date-format: long
abstract: "Motivated by the lack of findings on different mortality causes for different age groups, this paper examines the main causes of mortality among young adults and middle-aged adults in Switzerland using data from the WHO Mortality Database for the years 2010, 2015, and 2019. The paper found that the leading causes of death for young adults are self-harm and drug use disorders, while middle-aged adults are most affected by ischemic heart disease and trachea, bronchus, and lung cancers. Moreover, it is also found that one of the main mortality causes in both age groups is self-harm which highlights the importance of mental health."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

Therefore, in this paper, we will investigate the top mortality causes in young adults and middle-aged adults in Switzerland and discuss some trends and patterns in healthcare that were witnessed in the data. The estimand that we’re trying to estimate is the number of deaths by different mortality causes in young adults and middle-aged adults. We aim to provide empirical evidence on different causes of death in two age groups and investigate the reasons behind the differences in health issues in two age groups. This paper uses data obtained from the open database, WHO Mortality database, which exhibits information on mortality causes of deaths filtered by country, sex, year, and age group. The data used in this paper were the Mortality causes in Switzerland in the years 2010,2015, and 2019 as it gives a general idea of the mortality causes from 2010 to 2019, filtered by age group.

@@

## Clean Data
he data utilized for this paper is downloaded manually through the WHO Mortality Database. All the data analysis was done through R programming language (@citeR) with the aid of the following packages:rstanarm (@rstanarm), tidyverse (@tidyverse), knitr (@knitr), modelsummary (@modelsummary), ggplot2 (@ggplot2), dplyr (@dplyr), parameters (@Parameters), broom (@broom), broom.mixed (@broommixed).

@@

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults2
#| tbl-cap: "Relationship between causes of deaths and number of deaths in middle-aged adults" 
#| warning: false
#| tbl-pos: H
modelsummary(
  list(
    " Above 40" = cause_of_death_above40
))
```
```

>>>>

---
title: "Top Mortality causes in young adults and middle age adults in Switzerland"
subtitle: "Analyzing Switzerland Mortality causes data in 2010, 2015, adn 2019 from WHO database"
author: 
  - Thu Dong
thanks: "Code and data are available at: https://github.com/thu2912/Top-mortality-causes-of-age-groups-in-Switzerland.git"
date: today
date-format: long
abstract: "Motivated by the lack of findings on different mortality causes for different age groups, this paper examines the main causes of mortality among young adults and middle-aged adults in Switzerland using data from the WHO Mortality Database for the years 2010, 2015, and 2019. The paper found that the leading causes of death for young adults are self-harm and drug use disorders, while middle-aged adults are most affected by ischemic heart disease and trachea, bronchus, and lung cancers. Moreover, it is also found that one of the main mortality causes in both age groups is self-harm which highlights the importance of mental health."
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

@@

Therefore, in this paper, we will investigate the top mortality causes in young adults and middle-aged adults in Switzerland and discuss some trends and patterns in healthcare that were witnessed in the data. The estimand that we’re trying to estimate is the number of deaths by different mortality causes in young adults and middle-aged adults. We aim to provide empirical evidence on different causes of death in two age groups and investigate the reasons behind the differences in health issues in two age groups. This paper uses data obtained from the open database, WHO Mortality database, which exhibits information on mortality causes of deaths filtered by country, sex, year, and age group. The data used in this paper were the Mortality causes in Switzerland in the years 2010, 2015, and 2019 as it gives a general idea of the mortality causes from 2010 to 2019, filtered by age group.

@@

## Clean Data
he data utilized for this paper is downloaded manually through the WHO Mortality Database. All the data analysis was done through R programming language (@citeR) with the aid of the following packages:rstanarm (@rstanarm), tidyverse (@tidyverse), knitr (@knitr), modelsummary (@modelsummary), ggplot2 (@ggplot2), dplyr (@dplyr), parameters (@parameters), broom (@broom), broom.mixed (@broommixed).

@@

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults2
#| tbl-cap: "Relationship between causes of deaths and number of deaths in middle-aged adults" 
#| warning: false
#| tbl-pos: H
modelsummary(
  list(
    " Above 40" = cause_of_death_above40
))
```
